---
title: "Binomial Confidence Intervals for Sensitivity and Other Proportions"
output: rmarkdown::html_vignette
bibliography: REFERENCES.bib
vignette: >
  %\VignetteIndexEntry{Binomial Confidence Intervals for Sensitivity and Other Proportions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette describes the binomial confidence-interval methods implemented in `bstatUtils` and shows how to use `binom_ci()` and `binom_ci_all()` in practice.

The motivating example is diagnostic sensitivity, defined as the proportion of test-positive results among truly diseased subjects. The same methods apply directly to any single binomial proportion, including specificity, prevalence, response probability, or event risk.

## Overview

For `n` successes out of `N` observations, the estimated proportion is `n / N`. The package implements seven interval methods: Wald, Wilson, Clopper-Pearson, Agresti-Coull, Jeffreys, logit, and likelihood-ratio [@Wilson1927; @ClopperPearson1934; @AgrestiCoull1998; @BrownCaiDasGupta2001].

These methods differ mainly in small-sample performance, behavior near 0 or 1, and whether the interval is symmetric or asymmetric on the probability scale. The Wald interval is easy to compute but is widely regarded as a poor default for small samples or extreme proportions, whereas Wilson and Jeffreys are commonly recommended as stronger general-purpose choices [@BrownCaiDasGupta2001].

## Why several methods?

No single interval is best in every setting.

The Wald interval is simple, but it is well known to perform poorly for small samples or proportions near the boundaries [@BrownCaiDasGupta2001]. Wilson and Jeffreys are often stronger default choices, while Clopper-Pearson is commonly used when an exact binomial method is preferred [@BrownCaiDasGupta2001; @ClopperPearson1934]. Agresti-Coull is a simple adjusted approximation [@AgrestiCoull1998], and the logit and likelihood-ratio intervals provide asymmetric alternatives that can behave better near 0 or 1 [@BrownCaiDasGupta2001].

## Implemented methods

| Method | Main idea | Typical strengths | Main cautions |
|:--|:--|:--|:--|
| Wald | Normal approximation around `n / N` | Very simple and familiar | Often inaccurate for small `N` or extreme proportions [@BrownCaiDasGupta2001] |
| Wilson | Score-test inversion | Strong general-purpose frequentist default [@Wilson1927; @BrownCaiDasGupta2001] | Slightly less transparent algebraically |
| Clopper-Pearson | Exact binomial interval from beta quantiles | Valid at the boundaries, exact under the binomial model [@ClopperPearson1934] | Often conservative and wider than needed [@BrownCaiDasGupta2001] |
| Agresti-Coull | Adjusted Wald interval using pseudo-counts | Simple and usually better than Wald [@AgrestiCoull1998] | Still approximate |
| Jeffreys | Equal-tailed beta interval with Jeffreys prior | Good small-sample and boundary behavior [@BrownCaiDasGupta2001] | Bayesian derivation may not suit every reporting context |
| Logit | Wald interval on the logit scale, then back-transform | Asymmetric and remains within `(0, 1)` | Needs boundary correction at `n = 0` or `n = N` |
| Likelihood-ratio | Inverts the binomial likelihood-ratio test | Likelihood-based and often sensible near boundaries [@BrownCaiDasGupta2001] | More computationally intensive |

## Functions

The package provides two user-facing functions.

- `binom_ci(n, N, method = "wilson", conf.level = 0.95, cc = 0.5)` computes one interval for one method.
- `binom_ci_all(n, N, conf.level = 0.95, cc = 0.5)` computes all implemented methods for the same data.

The `method` argument accepts the following values:

```{r}
c(
  "wald",
  "wilson",
  "clopper-pearson",
  "agresti-coull",
  "jeffreys",
  "logit",
  "likelihood"
)
```

The `cc` argument is used only by the logit method when `n = 0` or `n = N`. In those boundary cases, the ordinary logit is undefined, so a continuity correction is used before transforming to the logit scale.

## Basic examples

A single Wilson interval:

```{r}
library(bstatUtils)

binom_ci(19, 20, method = "wilson")
```

A Clopper-Pearson interval for the same data:

```{r}
binom_ci(19, 20, method = "clopper-pearson")
```

All implemented methods at once:

```{r}
binom_ci_all(19, 20)
```

Each result is returned as a `data.frame` with the observed successes `n`, total `N`, point estimate `estimate`, method name, confidence level, and the lower and upper confidence limits.

## Boundary cases

Boundary cases deserve special attention because many interval methods behave differently when `n = 0` or `n = N` [@BrownCaiDasGupta2001].

For example, the logit interval uses the continuity correction `cc` only at the boundaries:

```{r}
binom_ci(0, 10, method = "logit")
binom_ci(10, 10, method = "logit")
```

The Clopper-Pearson interval handles the boundaries naturally:

```{r}
binom_ci(0, 10, method = "clopper-pearson")
binom_ci(10, 10, method = "clopper-pearson")
```

## Choosing a method

A practical default is Wilson when a frequentist interval is wanted and no special requirement points elsewhere [@BrownCaiDasGupta2001; @Wilson1927].

Jeffreys is a strong alternative when small-sample performance and boundary behavior are especially important [@BrownCaiDasGupta2001]. Clopper-Pearson is appropriate when an exact interval is required by convention, regulation, or reporting policy, keeping in mind that it is usually conservative [@ClopperPearson1934; @BrownCaiDasGupta2001]. Agresti-Coull is a simple adjusted approximation [@AgrestiCoull1998], while logit and likelihood-ratio intervals are useful when asymmetric limits are preferred.

For routine reporting, the Wald interval is usually best treated as a benchmark rather than a recommended default [@BrownCaiDasGupta2001].

## Input validation

The functions check that:

- `n` is a single non-missing integer-like numeric value,
- `N` is a single positive integer-like numeric value,
- `0 <= n <= N`,
- `conf.level` lies strictly between 0 and 1,
- `cc >= 0`.

Examples of invalid inputs:

```{r, error = TRUE}
binom_ci(11, 10, method = "wilson")
```

```{r, error = TRUE}
binom_ci(1.5, 10, method = "wilson")
```

## References {-}
